---
title: "What is your blogpost title? Add it here??"
subtitle: "Assignment 4 ETC5512"
author: "Pranav Sajeev - 35571381"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

## Data and Documenting

### 1. Research Question

What is the ideal length of a game that gives the maximum amount of satisfaction, by genre?

For each genre, what game lengths tend to receive the highest player ratings?\
Do different genres have different “ideal” lengths in terms of user satisfaction?

## 2. The data

The data is downloaded mainly from 2 sources:

1.   Data from the Steam store: <https://github.com/NewbieIndieGameDev/steam-insights>
2.  PlaymyData, which is a dataset that contains information about the amount of time taken to complete each game. It also contains other metrics. This data is taken from the website <https://huggingface.co/datasets/claudioDsi94/PlayMyData/tree/main>.

## 3. Processing the data

```{r}
#Loading Libraries
library(tidyverse)
library(readr)
library(stringr)
```

```{r}
# Reading in the data

steam_game_data=read_csv(("./data/games.csv"))
playmydata_pc_data=read_csv("./data/all_games_PC.csv")
#Reading in the steamspy insights data 
steamspy_data= read_csv("./data/steamspy_insights.csv")


```

```{r}
##Removing all demos from the data and NA values
steam_game_data = steam_game_data|>
  filter(!type=="demo")

steam_game_data = steam_game_data|>
  filter(!owners_range=="NA")
#Selecting the required variables from the steam dataset
steam_game_data=steam_game_data|>
  select(app_id,name)
#Removing the /N values from steamspy
  steamspy_data= steamspy_data|>
    filter(!genres=="\\N")|>
    filter(!owners_range=="\\N")
#Selecting only the required variables
  steamspy_data=steamspy_data|>
    select(app_id,owners_range,genres)
  
#Joining the raw game.csv to the steamspy like we did for categories and tags
steam_game_data=left_join(steam_game_data,steamspy_data, by = "app_id")
#Removing app_id as we don't need it going forward
steam_game_data = steam_game_data|> select(-app_id)


#Some names have  "™" so we will remove them to avoid problems when joining the datasets
steam_game_data$name = str_replace_all(steam_game_data$name, "™", "")

#Now cleaning the playmydata data

playmydata_pc_data=playmydata_pc_data|>
  select(-c(genres,id,platforms,summary,storyline,people_polled,))
#Removing NA values
playmydata_pc_data=playmydata_pc_data|>
  filter(!rating %in% c("Missing","nan"))
playmydata_pc_data=playmydata_pc_data|>
  filter(!main =="NaN")|>
  filter(!main== 0.00)

#Ratings are stored as characters, so changing them to double
playmydata_pc_data$rating = as.numeric(playmydata_pc_data$rating)
#Rounding the rating values
playmydata_pc_data$rating = round(playmydata_pc_data$rating,2)


#Joining the datasets
final_game_data = inner_join(playmydata_pc_data,steam_game_data, by="name")

# adjustments on the final dataset
#Separating the genres into different rows
final_game_data = final_game_data|> 
  separate_rows(genres, sep = ",\\s*") 

#Creating an average length variable based on completeion times
final_game_data_t = final_game_data|> 
  mutate(
  main = na_if(main, 0),
  extra = na_if(extra, 0),
  completionist = na_if(completionist, 0),
  avg_length = rowMeans(across(c(main, extra, completionist)), na.rm = TRUE)
)


```

### Remember

Please mention any additional files that you want the markers to review, eg. read me, meta data etc.

The github repository for this Quarto document is [here](https://github.com/Psaj0002/ETC5512-Assignment-2--35571381). Please note that the first commit is in a separate file as the file was later renamed.

<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

### What's in this section

Here is where you should write your blogpost! (Task 2)

### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)

2.  Motivation

    Winter break is almost upon us. Many have plans set for the holidays already; some plan to travel, some plan to work, some plan to take up some extra studies, the possibilities are endless. Personally, I want to sit and play as many video games as possible. However, time is limited and this leads to an important question: which games should I choose to play?

    Each game can take days to finish, given that I am taking enough breaks to rest, and there are so many different types out there, from small indie games to large open-world RPGs. So it is important to choose the games I play wisely to make sure I don't feel like I've started something I can't finish.

    This inspired me to investigate what the ideal length of a game for different genres are, in terms of both duration and satisfaction. Are players more satisfied with shorter or longer games in each genre? Can we identify “sweet spots” where games are most positively received?

3.  \

4.  Data\

5.  Analysis\

6.  Conclusions\

7.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

### Question 5

1.  the tags and categories were in a wide format. So each tag/category was displayed in it's own row. Insetad I made the dataset shorter by making all the tags/categories into a vector for each id.
2.  Some games in the Steam data had the ™ in the name, which led to problems while joining the datasets by name

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
