---
title: "What is your blogpost title? Add it here??"
subtitle: "Assignment 4 ETC5512"
author: "Who are you?? Add your name here"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

## Data and Documenting

### 1. Research Question

What is the ideal length of a game that gives the maximum amount of satisfaction, by genre?

For each genre, what game lengths tend to receive the highest player ratings?\
Do different genres have different “ideal” lengths in terms of user satisfaction?

## 2. The data

1.  the data is downloaded mainly from 2 sources: <https://github.com/NewbieIndieGameDev/steam-insights> for data from the Steam store.
2.  PlaymyData, which is a dataset that contains information about the amout of time taken to complete each game. It also contains other metrics. This data is taken from the website <https://huggingface.co/datasets/claudioDsi94/PlayMyData/tree/main>

## 3. Processing the data

```{r}
#Loading Libraries
library(tidyverse)
library(readr)
```

```{r}
# Reading in the data


playmydata_pc_data=read_csv("./data/all_games_PC.csv")

steam_game_data=read_csv(("./data/games.csv"))

steam_categories = read_csv("./data/categories.csv")

steam_tags = read_csv("./data/tags.csv")
```

```{r}
# Grouping the tags and category data
steam_tags_grouped = steam_tags |>
  group_by(app_id) |>
  summarise(tags = list(tag), .groups = "drop")

steam_categories_grouped = steam_categories|>
  group_by(app_id) |>
  summarise(categories= list(category), .groups = "drop")

# Adding the data to the main steam data
steam_game_data = steam_game_data |>
  left_join(steam_categories_grouped, by= "app_id")|>
  left_join(steam_tags_grouped, by="app_id")

#Creating the final dataset by joining the playmydata dataset
final_game_data = left_join(steam_game_data, playmydata_pc_data, by = "name")
```

### Remember

Please mention any additional files that you want the markers to review, eg. read me, meta data etc.

<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

### What's in this section

Here is where you should write your blogpost! (Task 2)

### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)

2.  Motivation

    Winter break is almost upon us. Many have plans set for the holidays already; some plan to travel, some plan to work, some plan to take up some extra studies, the possibilities are endless. Personally, I want to sit and play as many video games as possible. However, time is limited and this leads to an important question: which games should I choose to play?

    Each game can take days to finish, given that I am taking enough breaks to rest, and there are so many different types out there, from small indie games to large open-world RPGs. So it is important to choose the games I play wisely to make sure I don't feel like I've started something I can't finish.

    This inspired me to investigate what the ideal length of a game for different genres are, in terms of both duration and satisfaction. Are players more satisfied with shorter or longer games in each genre? Can we identify “sweet spots” where games are most positively received?

3.  \

4.  Data\

5.  Analysis\

6.  Conclusions\

7.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

### Question 5

1.  the tags and categories were in a wide format. So each tag/category was displayed in it's own row. Insetad I made the dataset shorter by making all the tags/categories into a vector for each id.
2.   Some games in the Steam data had the ™ in the name, which led to problems while joining the datasets by name

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
